---
title: R Factors
author: ''
date: '2021-12-15'
slug: []
categories: []
tags: 
  - r
  - ggplot
subtitle: ''
summary: 'Ordering factors using `fct_inorder` and a step forward in personal productivity using `rmarkdown::draft`.'
authors: []
lastmod: '2021-12-15T19:47:00-05:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
editor_options: 
  chunk_output_type: console
---

I had a realization that I spend an enormous amount of time trying to custom order categorical variables on visualizations. However, I realized I can make this *way* easier on myself by using `fct_inorder` and a left join.

```{r, warning=F, message=F}
library(tidyverse)

VALS <- tribble(
  # by using fct_inorder, the factor will be ordered in whatever
  # order the labels appear in the dataset
  ~Species, ~lab,
  'setosa', 'Setosa (a)',
  'virginica', 'Virginica (c)',
  'versicolor', 'Versicolor (b)',
) %>% mutate_at(c('lab'), ~fct_inorder(.x))
```

Now I join the `VALS` dataset with my source data and use the `lab` variable as the xaxis. Whenever I want to shuffle the values around, I can just rearrange them in the `VALS` dataset.
```{r, warning=F, message=F}
iris %>% 
  left_join(VALS, by=c('Species')) %>% 
  group_by(lab) %>% 
  summarise(avg = mean(Petal.Width)) %>% 
  ggplot(aes(lab, avg, fill=lab)) +
  geom_col()+
  geom_text(aes(label=round(avg, 2)), nudge_y=.05) +
  scale_fill_brewer(palette='Dark2')+
  scale_y_continuous(
    limits=c(0, 2.5),
    breaks=seq(0, 2.5, 1),
    expand = expansion(add=c(0, .1))
  )+
  labs(
    x='', y='Average Petal Width',
    title = 'Ordered Factors',
    fill = 'Category'
  )+
  ggthemes::theme_clean()+
  theme(
    axis.ticks = element_blank(),
    axis.title = element_text(size=13),
    axis.text = element_text(size=12),
    axis.text.y = element_blank(),
    legend.position = 'none'
  )
```

---- 

# Bonus R markdown template

My R projects at work usually have 1 main R Markdown notebook and then a `queries/` folder that contains queries I commonly use. I usually end up with a folder structure something like this:

- `.gitignore`
- `readme.md`
- `analysis.Rmd`
- `queries/`
  - `getdat1.sql`
  - `getdat2.sql`
  
Now, I *do* have an R Markdown template that I use for all my notebooks, but for my sql queries I would do this really annoying copy and paste routine every time I wanted to use a sql query.

1. Create the `analysis.Rmd` document using an R Markdown template that I created.
2. Do some work, then realize I need to pull data on `x`.
3. Copy the query that gets `x` from a centralized sql folder on my machine to my project folder. 
4. Do some more work, then realize I need to pull data on `y`. 
5. Copy the query that gets `y` from a centralized sql folder on my machine to my project folder. 
6. Go back to #2 and repeat.
7. Finish the project lol

[This chapter of R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/document-templates.html) goes into detail about how to set up a template in the first place and how to incorporate supporting files (my `queries/` folder). 

Adding the supporting files to my template was painless, but whenever I would create a new rmarkdown document from that template it would put all the newly generated files into a subfolder which was really annoying. My projects ended up looking like this:

- `.gitignore`
- `mytemplate/`
  - `analysis.Rmd`
  - `queries/`
    - `getdat1.sql`
    - `getdat2.sql`

To get around this behavior (updating `create_dir` in the template yaml does not change this behavior) I had to use the `rmarkdown::draft` function in the R console instead of using the RStudio gui to use the template. So instead of clicking through dropdowns I'd send the following in the console:

```
rmarkdown::draft(
  "tocreate.Rmd", 
  template = "name_of_tmplt", 
  package = "name_of_pkg"
)
```

and now the directory structure is preserved.

I created a snippet called "begin" that is simply the above code, that way whenever I want to create a new project I can just type `begin` and get started. 

Maybe a video explanation would be clearer...