---
title: Visualizing Spatial Data
author: ''
date: '2021-11-28'
slug: []
categories: []
tags: []
subtitle: ''
summary: 'with `leaflet` and `tidygeocoder`'
authors: []
lastmod: '2021-11-28T15:55:19-05:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
editor_options: 
  chunk_output_type: console
---

For some reason I've always been afraid of working with spatial data, but `leaflet` and `tidygeocoder` have made all my worries go away! Here's an example using some addresses in New York.

```{r, warning=F, message=F}
library(tidyverse)
library(leaflet)
library(tidygeocoder)

knitr::opts_chunk$set(warning = F, message = F)

locations <- tribble(
  ~addr, ~label,
  '26 Broadway, New York, NY 10004', 'The Bull (kinda)',
  '11 Wall St, New York, NY 10005', 'NYSE',
  '90 Church St Fl 1, New York, NY 10007', 'World Trade Center',
  '205 E Houston St, New York, NY 10002', 'Katz Deli',
  'Washington Square, New York, NY 10012', 'Washington Square Park'
) %>% 
  # Geocode an address
  # First try the census api, but if that doesn't yield any results
  #   try the Nominatim api
  geocode_combine(
    queries = list(
      list(method='census'),
      list(method='osm')
    ),
    global_params = list(address = 'addr')
  )
```

Once I've geocoded the addresses into latitude and longitude, I can use `leaflet` to visualize the results.
```{r}
mymap <- leaflet() %>% 
  addTiles() %>% 
  addMarkers(
    data = locations,
    label = ~label
  )  

mymap
```

<br>

There are many different map styles [(aka "Third Party Tiles")](https://rstudio.github.io/leaflet/basemaps.html) to choose from and they can be implemented easily using `addProviderTiles`.

The map below is going to look identical to the one above it, which is odd because it works properly in `R Studio` but once I knit the document the style vanishes...use your imagination!!

```{r}
mymap %>% 
  addProviderTiles(providers$Stamen.Toner)
```
