---
title: Data Wrangling Doodles
author: R package build
date: '2021-09-18'
slug: []
categories: []
tags: []
subtitle: ''
summary: 'Using `flextable`, `janitor`, `waldo`, and `forcats`'
authors: []
lastmod: '2021-09-18T09:41:50-04:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
editor_options: 
  chunk_output_type: console
draft: false
---

This post shows 3 data cleaning things I've been doing quite a bit lately:

1. Building and formatting tables
2. Comparing dataset columns / types
3. Combining categories

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(flextable)
library(janitor)
library(waldo)
```

---

`janitor` and `flextable` are really great for building and formatting tables:

```{r message=FALSE, warning=FALSE}
mtcars %>%
  tibble() %>%
  janitor::tabyl(cyl) %>%
  janitor::adorn_pct_formatting(digits=0) %>%
  flextable() %>%
  theme_booktabs() %>%
  set_header_labels(
    cyl = 'Cylinder',
    n = '# Obs',
    percent = 'Percent'
  ) %>%
  add_header_row(values=c('Summary Table'), colwidths = c(3)) %>%
  align(align='center', part='all') %>%
  bg(bg = 'blue', part='header') %>%
  color(color='white', part='header')
```

---

The `compare` function from `waldo` makes joining and concatenating data that isn't exactly the same less painful: 

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
base <- tribble(
  ~name, ~height,
  'tony', '15',
  'bob', '13'
)

aug <- tribble(
  ~NAME, ~height,
  'tony', 15,
  'bob', 13
)

base %>% 
  flextable() %>% 
  add_header_lines(values='base') %>%
  align(align='center', part='all') %>%
  flextable_to_rmd()

aug %>% 
  flextable() %>% 
  add_header_lines(values='aug') %>%
  align(align='center', part='all') %>%
  flextable_to_rmd()
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
base %>%
  waldo::compare(aug,
          x_arg='base',
          y_arg='aug'
          ) 
```

There is also a function in `janitor` that does something similar
```{r}
janitor::compare_df_cols(
  base,
  aug
) %>% 
  flextable() %>%
  align(align='center', part='all')
```

---

Combining factors with `forcats` is really handy when the categories are at different levels. For example, combining the 'Cats' and 'Dogs' categories in one data source to match the 'Animals' category in another.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
a <- tribble(
  ~Category, ~pct, ~dat_src,
  'Cats', .15,   'A',
  'Dogs', .40,   'A',
  'Humans', .45, 'A'
)

b <- tribble(
  ~Category, ~pct, ~dat_src,
  'Animals', .25, 'B',
  # Are humans animals? thinkingemoji.jpg
  'Humans', .75, 'B'
)
```

I can go from this:

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
a %>% 
  flextable() %>% 
  bg(i = ~ Category %in% c('Cats', 'Dogs'), j=~ Category, bg='#edd695') %>%
  align(align = 'center', part='all') %>%
  flextable_to_rmd()
  
b %>% 
  flextable() %>% 
  bg(i= ~ Category == 'Animals', j=~Category, bg='#edd695') %>%
  align(align = 'center', part='all') %>%
  flextable_to_rmd()
```

to this

```{r, echo=FALSE, message=FALSE, warning=FALSE}
a %>%
  mutate_at(c('Category'), as_factor) %>%
  mutate_at(c('Category'), 
            ~ fct_collapse(.x, Animals = c('Cats', 'Dogs'))) %>%
  group_by(Category, dat_src) %>%
  summarise(pct = sum(pct)) %>% 
  ungroup() %>%
  bind_rows(b) %>%
  pivot_wider(
    names_from=dat_src,
    values_from=pct
  ) %>%
  flextable() %>%
  align(align = 'center', part='all') 
```

by doing this:
```r 
a %>%
  mutate_at(c('Category'), as_factor) %>%
  mutate_at(c('Category'), 
            ~ fct_collapse(.x, Animals = c('Cats', 'Dogs'))) %>%
  group_by(Category, dat_src) %>%
  summarise(pct = sum(pct)) %>% 
  ungroup() %>%
  bind_rows(b) %>%
  pivot_wider(
    names_from=dat_src,
    values_from=pct
  ) %>%
  flextable()
```